### 2023-10-11 main.tex
,

\label{rhox}

@@

Consider two Markov processes, $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 

@@

The duality definition can also be formulated as a relation between the generators, provided some minimal technical conditions are satisfied \cite{jansen2014notion}. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$

@@

when the state spaces of the dual processes is countable, the generators of the processes and the duality function can be represented as matrices with elements $\mathcal{L}(\eta,\eta')$, $\widetilde{\mathcal{L}}(\xi,\xi')$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta'\in\Omega$ and $\xi,\xi'\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
    \sum_{\eta'\in\,\Omega}\mathcal{L}(\eta,\eta')D(\eta',\xi)=\sum_{\xi'\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi')D(\eta,\xi')

@@

where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}^{\,T}$. Working in terms of the Hamiltonian operators the duality relation reads 

@@

\subsection{Duality for the open multi-species stirring process}\label{statementDualitySubsection}
In this section we formulate duality for the multi-species stirring process $(\bm{n}(t))_{t\geq 0}$ with open boundaries, defined by the generator \eqref{Generator}.

@@

This means that to each site $x\in V$ we associate an ``extra-site'' via a bijection $u:V\to V$. We denote as $u(x)$ the extra-site asspciated to $x\in V$ . The configuration space of the dual process is the enlarged state space

@@

    \widetilde{\Omega}= \bigotimes_{x\in V} \widetilde{\Omega}_{x}\ = \bigotimes_{x\in V} (\Omega_{x}\times \mathbb{N}_{0}^{N})\dt
Note that on the extra-site we allow an unbounded number of particles. Thus
dual particle will accumulate in these extra-sites in the course of time. 
We write the configurations $\bm{\xi} \in \widetilde\Omega$  as

@@

We define the duality function of the open multi-species stirring process as
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\,\right)
\end{equation}
where we recall the notation (see \eqref{rhox}) for the \textit{density} of the species $a\in \{1,\ldots,N-1\}$ imposed by the reservoir at site $x\in V$:
\begin{equation}
\rho_{a}^{x}=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}\dt
\end{equation}
The dual process is defined by its Markov generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=1}^{N-1}\xi_{a}^{x}\left(f(\bm{\xi}-\bm{\delta}_{a}^{x}+\bm{\delta}_{N}^{x}+\bm{\delta}_{a}^{u(x)})-f(\bm{\xi})\right)\dt
\end{equation}
\newline
Therefore the dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual Markov generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{1,\ldots,N-1\}$ at site $x$ with a particle of type $N$ and creates a particle of the same type $a$ at the extra-site $u(x)$. This last transition is performed with rate $|\alpha^{x}|\xi_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{1,\ldots,N-1\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\textbf{Remark}: in the reversible situation, i.e. when $\forall x\in V$ we have $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigotimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=1}^{N-1}\left(\rho_{a}\right)^{\sum_{x\in V}\xi_{a}^{x}+\sum_{x\in V}\xi_{a}^{u(x)}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}\dt
\end{equation}

\subsection{Proof of duality}
To prove duality between $(\bm{n}(t))_{t\geq 0}$ and $(\bm{\xi}(t))_{t\geq 0}$ we  show that \eqref{DualityRelation} is fulfilled.  To show this, we will use the Hamiltonians and their Lie algebraic description. Indeed, in this formalism the proof 
reduces to finding symmetries of the generator (for the bulk duality) and group like transformations (for the ``boundary duality'').

To a configuration $\bm{\xi}$ of the configuration space  \eqref{dualStateSpace} of a dual process we associate the vector
\begin{equation}
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes |\xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}\rangle\right)\dt
\end{equation}
The Hamiltonian of the dual process reads
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\mathcal{H}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}
where $\mathcal{H}_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=|\alpha^{x}|\sum_{a=1}^{N-1}\left((A^{\dagger})_{a}^{u(x)}\,E_{Na}^{x}-E_{aa}^{x}\right)\dt
\end{equation}
Here we introduced bosonic creation operator $A^{\dagger}$ acting on the left-hand-side as $\langle q|A^{\dagger}=\langle q+1|$ on a generic vector $\langle q|$ with $q\in \mathbb{N}_{0}$, so that in \eqref{siteDualHamiltonian} 
$(A^{\dagger})_{a}^{u(x)}$ denotes a copy of $A^{\dagger}$ acting on the extra-site $u(x)$ and on the species $a\in\{1,\ldots,N-1\}$. \\
We will show below that the Hamiltonians \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} are dual in the sense of \eqref{DualityRelation}. From an algebraic point of view, the duality matrix $D$ \eqref{dualityElements} is described as 
\begin{equation}\label{dualityMatrix}
    D=\prod_{x\in V}d_{x}\otimes \dd \;.
\end{equation}
Here
\begin{equation}\label{bulkElementDualityMatrix}
d_{x}=R_{x}\exp{(E^{x})}
\end{equation}
with 
\begin{equation}\label{Revmatrix}
    R_{x}=\sum_{n^{x}\in\Omega_{x}}\frac{\prod_{a=1}^{N}n_{a}^{x}!}{\nu!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle\langle n_{1}^{x},\ldots,n_{N}^{x}|
\end{equation}
and
\begin{equation}\label{EquationEx}
E^{x}=\sum_{a=1}^{N-1}E_{aN}^{x}\,.
\end{equation}
Furthermore
\begin{equation}\label{dualityMatrix2}
\dd=\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}|\dt
\end{equation}
We observe that the matrix $R_{x}$ is diagonal. Its elements are related to the inverse of the weights of the reversible measure \eqref{reversibleMeasure}. In particular, to obtain these elements, we have considered the weights of \eqref{reversibleMeasure} when all the parameters $\rho_{a}=\frac{1}{N}$. Then, the constant $\left(\frac{1}{N}\right)^{\nu}$ has been neglected, since it does not change the duality relation. This $R_{x}$ is called the ``cheap'' duality matrix and is associated to reversibility (see \cite{giardina2009duality}). \\

Since \eqref{dualityMatrix} is product over sites, proving \eqref{DualityRelation} is equivalent to showing that 
\begin{equation}\label{edgeDualRealtion}
    \mathcal{H}_{x,y}^{T}D=D\mathcal{H}_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\end{equation}
and 
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
\end{equation}
We perform the proof of duality in three steps: first we will show that matrix \eqref{dualityMatrix} has elements \eqref{dualityElements}; second we will prove the bulk duality \eqref{edgeDualRealtion}; finally we will show the boundary duality \eqref{siteDualRelation}. 
As a preliminary result we note that 
\begin{equation}\label{transpositionPropertyR}
(E_{ba}^{x})^{T}=R_{x}E_{ab}^{x}R_{x}^{-1}\qquad \forall x\in V
\end{equation}
that follows immediately from the definition of $E_{ab}$. 
\begin{comment}

@@

\paragraph{Elements of the duality matrix.}
Consider the matrix $D$ defined in \eqref{dualityMatrix}.
We aim to show that its matrix elements coincide with duality function \eqref{dualityElements}, i.e.

@@

\langle \bm{n}|D|\bm{\xi}\rangle=D(\bm{n},\bm{\xi})\qquad   \forall \bm{n}\in \Omega,\quad \bm{\xi}\in \widetilde{\Omega}\,.

@@

where {\color{red} we used the orthogonality relation \eqref{ortho}}. 

@@

where $H^{'}= S_{1} H S_{1}^{-1} $ has both boundary terms in a triangular form, and $H^{''} = S_{2}H' S_{2}^{-1}$ has the left boundary in a triangular form and the right boundary is diagonal. Using these transformations the commutators \eqref{bulk} and the action on the boundary vectors \eqref{leftBoundary}, \eqref{rightBoundary} become simpler. As a consequence, the steady state $|\Psi^{''}\rangle$ of $H^{''}$ can be written explicitly. Finally, reversing the transformations we retrieve $|\Psi\rangle$.

@@

    S_{1}=\prod_{x=1}^{L}\mathcal{S}_{1}^{x}\dt

@@

S_{1}=\prod_{x=1}^{L}d_{x}^{T}

@@

   H^{'}=H^{'}_{left}+H_{bulk}+H^{'}_{right}

@@

	|n^{x}\rangle =|\tau_{x}\rangle \qquad \text{and}\qquad|\bm{n}\rangle=|\bm{\tau}\rangle
with the meaning explained above. It will be useful to introduce also $\bm{\tau}=(\tau_{x})_{x\in \{1,\ldots,L\}}$. \\ \\

@@

	|V \rangle\nonumber&=\langle V^{*}|\widetilde{X}_{N}^{\sum_{i=1}^{L}n_{N}^{i}}\prod_{x=1}^{L}\prod_{a=1}^{N-1}\left(\lambda_{a}+\lambda_{a}\sum_{j=x}^{L}n_{N}^{j}\right)^{n_{a}^{x}}|V\rangle\nonumber

@@

To determine the state we now need to compute $\langle V^{*}|\widetilde{X}_{N}^{\sum_{i=1}^{L}n_{N}^{i}}|V\rangle$. Calling $\sum_{i=1}^{L}n_{N}^{i}=n_{N}$ and using \eqref{commLEFT} for $a\in \{1,\ldots,N-1\}$ we obtain 

@@

	|V \rangle=\Gamma(2+n_{N})\prod_{x=1}^{L}\prod_{a=1}^{N-1}\left(\lambda_{a}\right)^{n_{a}^{x}}\left(1+\sum_{j=x}^{L}n_{N}^{j}\right)^{n_{a}^{x}}

@@

\paragraph{Proof of formula \eqref{ABS-vect}} Using \eqref{S2-Inverse} we have $|\Psi^{'}\rangle=S_{2}^{-1}|\Psi^{''}\rangle$, thus we only need to show how the transformation $S_{2}^{-1}$ acts on the vector $|\Psi^{''}\rangle$. Here it is more convenient to consider the matrix $\mathcal{S}_{2}^{-1}$ defined in \eqref{S2Gen}, acting on a single site, and then make explicitly the tensor product over the chain. Using the exponential series, the matrix $\mathcal{S}_{2}$ is given by 
   \mathcal{S}_{2}^{-1}=\mathbbm{1}+\sum_{a=1}^{N-1}\beta_{a}e_{aN}\co
which action on the occupation variable of a site is 
  \mathcal{S}_{2}^{-1} |\mu\rangle= |\mu\rangle+\delta_{N,\mu}\sum_{a=1}^{N-1}\beta_a|a\rangle\dt
Therefore, taking the tensor product over the chain we obtain 
  \left(\mathcal{S}_{2}^{-1}\right)^{\otimes L} |\bm{\mu}\rangle=\left( |\mu_1\rangle+\delta_{N,\mu_1}\sum_{a=1}^{N-1}\beta_a|a\rangle\right)\otimes\ldots\otimes \left( |\mu_L\rangle+\delta_{N,\mu_L}\sum_{a=1}^{N-1}\beta_a|a\rangle\right)
By projecting over a vector $\langle \bm{\tau}|$ we have 
\begin{equation}\label{goodTP-S2}
  \langle \bm{\tau}|\left(\mathcal{S}_{2}^{-1}\right)^{\otimes L} |\bm{\mu}\rangle=\prod_{i=1}^L\left[ \delta_{\tau_i,\mu_i}+\delta_{N,\mu_i}\beta_{\tau_i} (1-\delta_{\tau_i,N})\right]
		N\qquad &\text{if}\quad x\in\{q_{1},\ldots,q_{i}\}\\
This is a new configuration where the occupations $\tau_{q_{1}},\ldots,\tau_{q_{i}}$ at sites $q_{1},\ldots,q_{i}$ are replaced with $N$'s, while the occupations on the other sites are left untouched. The dependence of $\varphi_{x}^{i}(q_{1},\ldots,q_{i})$ with respect to $q_{1},\ldots,q_{i}$ is usually understood for the sake of notation. By considering a fixed $\langle\bm{\tau}|$, and by using \eqref{goodTP-S2} we obtain 
\begin{equation}\label{ABS_intermediate}
		\Psi^{'}(\bm{\tau})&=\langle \bm{\tau}|(\mathcal{S}_{2}^{-1})^{\otimes L}|\Psi^{''}\rangle\\&= \Psi^{''}(\bm{\tau})+\sum_{q_{1}= 1}^{L}(1-\delta_{\tau_{q_{1}},N})\Psi^{''}(\varphi_{1}^{1},\ldots,\varphi_{L}^{1})\beta_{q_{1}^{1}}\\&+
		\sum_{q_{1},q_{2}= 1}^{L}(1-\delta_{\tau_{q_{1}},N})(1-\delta_{\tau_{q_{2}},N})\Psi^{''}(\varphi_{1}^{2},\ldots,\varphi_{L}^{2})\beta_{a_{1}}\beta_{a_{2}}\\&+\ldots+\Psi^{''}(N,\ldots,N)\left(\prod_{x=1}^{L}(1-\delta_{\tau_{x},N})\right)\prod_{x=1}^{L}\beta_{\tau_{x}}\dt%^{1-\delta_{\tau_{x},N}}
For a fixed configuration $\bm{\mu}=(\mu_{1},\ldots,\mu_{L})$, by using equation \eqref{ResulsBasis},we obtain 
	\Psi^{''}(\bm{\mu})=\frac{\Gamma(2+\sum_{x=1}^{L}\delta_{\mu_{x},N})}{\Gamma(L+2)}\prod_{x=1}^{L}\left(\lambda_{\mu_{x}}\left(2+L-x-\sum_{j=x}^{L}(1-\delta_{\mu_{j},N})\right)\right)^{1-\delta_{\mu_{x},N}}\co
\end{equation} where we have used the fact that 
	\left(1+\sum_{j=x}^{L}\delta_{\mu_{j},N}\right) =\left(2+L-y-\sum_{j=x}^{L}(1-\delta_{\mu_{x},N})\right)\co
Introducing $c_{x}$ defined as 
	c_{x}(\mu_{x})=(1-\delta_{\mu_{x},N}) \qquad \forall x\in \{1,\ldots L\}\co

>>>>

@@

Consider two Markov processes $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 

@@

The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$

@@

when the state spaces of the dual processes is finite, the generators and the duality function can be represented as matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})

@@

where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working in terms of the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads 

@@

\subsection{Duality for the multi-species stirring process}\label{statementDualitySubsection}
In this section we formulate duality for the multi-species stirring process $(\bm{n}(t))_{t\geq 0}$, defined by the generator \eqref{Generator}.

@@

This means that to each site $x\in V$ we associate an ``extra-site'', denoted $u(x)$. The configuration space of the dual process is the enlarged state space

@@

    \widetilde{\Omega}= \bigtimes_{x\in V} \widetilde{\Omega}_{x}\ = \bigtimes_{x\in V} (\Omega_{x}\times \mathbb{N}_{0}^{N})\dt
Dual particle will accumulate in these extra-sites in the course of time. We write the configurations $\bm{\xi} \in \widetilde\Omega$  as

@@

The duality function is given by
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\,\right)
\end{equation}
where we denote the \textit{density} of the species $a\in \{1,\ldots,N-1\}$ imposed by the reservoir at site $x\in V$  by 
\begin{equation}
\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}\dt
\end{equation}
The dual process is defined by its Markov generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=1}^{N-1}\xi_{a}^{x}\left(f(\bm{\xi}-\bm{\delta}_{a}^{x}+\bm{\delta}_{N}^{x}+\bm{\delta}_{a}^{u(x)})-f(\bm{\xi})\right)\dt
\end{equation}
\newline
Therefore the dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual Markov generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{1,\ldots,N-1\}$ at site $x$ with a particle of type $N$ and creates a particle of the same type $a$ at the extra-site $u(x)$. This last transition is performed with rate $|\alpha^{x}|\xi_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{1,\ldots,N-1\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\textbf{Remark}: in the reversible situation, i.e. when $\forall x\in V$ we have $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigtimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=1}^{N-1}\left(\rho_{a}\right)^{\sum_{x\in V}\xi_{a}^{x}+\sum_{x\in V}\xi_{a}^{u(x)}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}\dt
\end{equation}

\subsection{Proof of duality}
To prove duality between $(\bm{n}(t))_{t\geq 0}$ and $(\bm{\xi}(t))_{t\geq 0}$ we  show that \eqref{dualityIntertwines} is fulfilled.  To show this, we will use the Hamiltonians (linked with the Markov generators by \eqref{Hamiltonian-Generator}) and their Lie algebraic description. Indeed, in this formalism the proof reduces to finding symmetries and group like transformation of generators of the Lie algebra.\\
To a configuration $\bm{\xi}$ of the configuration space  \eqref{dualStateSpace} of a dual process we associate the vector
\begin{equation}
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes |\xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}\rangle\right)\dt
\end{equation}
The Hamiltonian of the dual process reads
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\mathcal{H}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}
where $\mathcal{H}_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=|\alpha^{x}|\sum_{a=1}^{N-1}\left((a^{\dagger})_{a}^{u(x)}\,E_{Na}^{x}-E_{aa}^{x}\right)\dt
\end{equation}
Here we introduced bosonic creation operator $a^{\dagger}$ acting on the left-hand-side as $\langle q|a^{\dagger}=\langle q+1|$ on a generic vector $\langle q|$ with $q\in \mathbb{N}_{0}$, so that in \eqref{siteDualHamiltonian} 
$(a^{\dagger})_{a}^{u(x)}$ denotes a copy of $a^{\dagger}$ acting on the extra-site $u(x)$ and on the species $a\in\{1,\ldots,N-1\}$. \\
We will show below that the Hamiltonians \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} are dual in the sense of \eqref{DualityRelation}. The duality matrix $D$ is described as 
\begin{equation}\label{dualityMatrix}
    D=\prod_{x\in V}d_{x}\otimes \dd
\end{equation}
where
\begin{equation}\label{bulkElementDualityMatrix}
d_{x}=R_{x}\exp{(E^{x})}
\end{equation}
with 
\begin{equation}\label{EquationEx}
E^{x}=\sum_{a=1}^{N-1}E_{aN}^{x}
\end{equation}
and
\begin{equation}\label{Revmatrix}
    R_{x}=\sum_{n^{x}\in\Omega_{x}}\frac{\prod_{a=1}^{N}n_{a}^{x}!}{\nu!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle\langle n_{1}^{x},\ldots,n_{N}^{x}|
\end{equation}
and where 
\begin{equation}\label{dualityMatrix2}
\dd=\sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}|\dt
\end{equation}
The matrix $R_{x}$ is diagonal. Its elements are related to the inverse of the weights of the reversible measure \eqref{reversibleMeasure}. In particular, to obtain these elements, we have considered the weights of \eqref{reversibleMeasure} when all the parameters $p_{i}=\frac{1}{N}$. Then, the constant $\left(\frac{1}{N}\right)^{\nu}$ has been neglected, since it does not change the duality relation. This $R_{x}$ is called the ``cheap'' duality matrix (see \cite{giardina2009duality}). \\
Since \eqref{dualityMatrix} is product over sites, proving \eqref{DualityRelation} is equivalent to showing that 
\begin{equation}\label{edgeDualRealtion}
    \mathcal{H}_{x,y}^{T}D=D\mathcal{H}_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\end{equation}
and 
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
\end{equation}
We perform the proof of duality in three steps: first we will show that matrix \eqref{dualityMatrix} has elements \eqref{dualityElements}; second we will prove \eqref{edgeDualRealtion}; finally we will show \eqref{siteDualRelation}. 
As a preliminary result we note that 
\begin{equation}\label{transpositionPropertyR}
(E_{ba}^{x})^{T}=R_{x}E_{ab}^{x}R_{x}^{-1}\qquad \forall x\in V
\end{equation}
that follows immediately from the definition of $E_{ab}$. 
\begin{comment}

@@

\paragraph{Elements of the duality matrix.}We aim to show that 

@@

\langle \bm{n}|D|\bm{\xi}\rangle=D(\bm{n},\bm{\xi})\qquad   \forall \bm{n}\in \Omega,\quad \bm{\xi}\in \widetilde{\Omega}
with $D(\bm{n},\bm{\xi})$ defined in \eqref{dualityElements}. 

@@

where we used the orthogonality relation \eqref{ortho}. 

@@

where $H^{'}= S_{1} H S_{1}^{-1} $ has both boundary terms in a triangular form, and $H^{''} = S_{2}H' S_{2}^{-1}$ has the left boundary in a triangular form and the right boundary is diagonal. Using these transformations the commutators \eqref{bulk} and the action on the boundary vectors \eqref{leftBoundary}, \eqref{rightBoundary} become simpler. As a consequence, the steady state $|\Psi^{''}\rangle$ of $H^{''}$ can be written explicitly. Finally, reversing the transformations we can retrieve $|\Psi\rangle$.

@@

    S_{1}=\prod_{x=1}^{L}\mathcal{S}_{1}^{x}

@@

S_{1}=\prod_{x=1}^{L}d_{x}^{T}\dt

@@

   H^{'}:=H^{'}_{left}+H_{bulk}+H^{'}_{right}

@@

	|n^{x}\rangle =\tau_{x}\qquad \text{and}\qquad|\bm{n}\rangle=|\bm{\tau}\rangle
with the meaning explained above. It will be useful to introduce also $\bm{\tau}=(\tau_{x})_{x\in \{1,\ldots,L\}}$. \\

@@

	|V \rangle\nonumber&=\langle V^{*}|\widetilde{X}_{N}^{\sum_{i=1}^{L}n_{N}^{x}}\prod_{x=1}^{L}\prod_{a=1}^{N-1}\left(\lambda_{a}+\lambda_{a}\sum_{j=x}^{L}n_{N}^{j}\right)^{n_{a}^{x}}|V\rangle\nonumber

@@

To determine the state we now need to compute $\langle V^{*}|\widetilde{X}_{N}^{\sum_{i=1}^{L}n_{N}^{i}}|V\rangle$. Calling $\sum_{x=1}^{L}n_{N}^{x}=n_{N}$ and using \eqref{commLEFT} for $a\in \{1,\ldots,N-1\}$ we obtain 

@@

	|V \rangle=\frac{\Gamma(2+n_{N})}{\Gamma(2)}\prod_{x=1}^{L}\prod_{a=1}^{N-1}\left(\lambda_{a}\right)^{n_{a}^{x}}\left(1+\sum_{j=x}^{L}n_{N}^{j}\right)^{n_{a}^{x}}

@@

\paragraph{Proof of formula \eqref{ABS-vect}} Using \eqref{S2-Inverse} we have $|\Psi^{'}\rangle=S_{2}^{-1}|\Psi^{''}\rangle$, thus we only need to show how the transformation $S_{2}^{-1}$ acts on the vector $|\Psi^{''}\rangle$. Here it is more convenient to consider the matrix $\mathcal{S}_{2}^{-1}$ defined in \eqref{S2Gen}, acting on a single site, and then make explicitly the tensor product over the chain. Using the exponential series, the matrix $\mathcal{S}_{2}$ is given by 
\begin{equation}
   \mathcal{S}_{2}^{-1}=\mathbbm{1}+\sum_{a=1}^{N-1}\beta_{a}e_{aN}\co
\end{equation}
which action on the occupation variable of a site is 
\begin{equation}
  \mathcal{S}_{2}^{-1} |\mu\rangle= |\mu\rangle+\delta_{N,\mu}\sum_{a=1}^{N-1}\beta_a|a\rangle\dt
\end{equation} 
Therefore, taking the tensor product over the chain we obtain 
\begin{equation}
  \left(\mathcal{S}_{2}^{-1}\right)^{\otimes L} |\bm{\mu}\rangle=\left( |\mu_1\rangle+\delta_{N,\mu_1}\sum_{a=1}^{N-1}\beta_a|a\rangle\right)\otimes\ldots\otimes \left( |\mu_L\rangle+\delta_{N,\mu_L}\sum_{a=1}^{N-1}\beta_a|a\rangle\right)
\end{equation} 
By projecting over a vector $\langle \bm{\tau}|$ we have 
\begin{equation}\label{goodTP-S2}
  \langle \bm{\tau}|\left(\mathcal{S}_{2}^{-1}\right)^{\otimes L} |\bm{\mu}\rangle=\prod_{i=1}^L\left[ \delta_{\tau_i,\mu_i}+\delta_{N,\mu_i}\beta_{\tau_i} (1-\delta_{\tau_i,N})\right]
\end{equation} 
For a fixed configuration $\bm{\mu}=(\mu_{1},\ldots,\mu_{L})$, using equation \eqref{ResulsBasis} and observing that 
\begin{equation}
	\left(1+\sum_{j=y}^{L}\delta_{\mu_{j},N}\right) =\left(2+L-y-\sum_{j=y}^{L}(1-\delta_{\mu_{y},N})\right)
\end{equation}
we obtain 
\begin{equation}
	\Psi^{''}(\bm{\mu})=\frac{\Gamma(2+\sum_{x=1}^{L}\delta_{\mu_{x},N})}{\Gamma(L+2)}\prod_{x=1}^{L}\left(\lambda_{\mu_{x}}\left(2+L-x-\sum_{j=x}^{L}(1-\delta_{\mu_{j},N})\right)\right)^{1-\delta_{\mu_{x},N}}\dt
\end{equation}
Introducing $c_{x}$ defined as 
\begin{equation}
	c_{x}(\mu_{x})=(1-\delta_{\mu_{x},N}) \qquad \forall x\in \{1,\ldots L\}
\end{equation}
we obtain 
\begin{equation}\label{coefficients-psiSec}
	\Psi^{''}(\bm{\mu})=\frac{\Gamma(2+L-\sum_{x=1}^{L}c_{x})}{\Gamma(L+2)}\prod_{x=1}^{L}\left(\lambda_{\mu_{y}}\left(2+L-x-\sum_{j=x}^{L}c_{j}\right)\right)^{c_{x}}
\end{equation}
where, for the sake of notation we left understood the dependence on $\mu_{x}$ of $c_{x}$.
\begin{comment}
We introduce the following notation
\begin{equation}\label{notation1}
    e_{ab}=e_{a}^{b}=\left(\delta_{ac}\delta_{bd}\right)_{c,d\in\{1,\ldots,N\}}\qquad \forall a,b\in \{1,\ldots,N\}
\end{equation}
and 
\begin{equation}\label{notation2}
    e_{ab}\otimes e_{cd}=e_{a,c}^{b,d}=\left(\delta_{(a,b),(e,f)}\delta_{(c,d),(g,h)}\right)_{e,f,g,h\in \{1,\ldots,N\}}\qquad \forall a,b,c,d\in \{1,\ldots,N\}
    \end{equation}
    that can be generalized for $e_{\tau_{1},\ldots,\tau_{L}}^{q_{1},\ldots,q_{L}}$ with $\tau_{1},\ldots,\tau_{L},q_{1},\ldots,q_{L}\in \{1,\ldots,N\}$. With this notation we have that
